---
title: Experimental evidence for the importance of multi-trophicstructure in regulating species persistence
author: "I. Bartomeus, S. Saavedra, R. Rohr, O. Godoy"
output:
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

This code allows reproducing all results and raw figures from the paper. We start by loading the data and functions needed.

```{r data and functions, message=FALSE, warning=FALSE}
set.seed(1111) #to ensure reproducibility
source('functions/toolbox.R') #set of tools and functions
source('functions/alphantw.R') #function to plot aphas ntw
source('functions/feasibility_domain_simplex.R') #funcion to plot triangles
load(file = "data/matrices/plants.RData") #plant matrices as calculayted in plant_parameters.R
load(file = "data/matrices/polinators.RData") #pollinator matrices directly from raw data.
load(file = "data/matrices/betas.Rdata") #Beta matrices ordered as required for the functions 
if(!require(ggtern)) {install.packages("ggtern"); library(ggtern)}
if(!require(deSolve)) {install.packages("deSolve"); library(deSolve)}
if(!require(matrixcalc)) {install.packages("matrixcalc"); library(matrixcalc)}
```

First we ask if plant coexistence within their trophic level changes when considering pollintors.

```{r plants, message=FALSE, warning=FALSE}
#Calculate Omegas (feasability domain) for PLANTS ONLY:
C <- diag(c(1,1,1), 3) # constraints -> assume that r can be positive only
### loops are useful to eliminate numerical instabilities and allow the value to converge
print(Omega_overlap(alpha,C)*2)
print(Omega_overlap(alpha_no_link,C)*2)
print(Omega_overlap(alpha_link,C)*2)

#PLOT Fig 2
#A) Ntw
mk_graph_3sp(alphamat = alpha, rs = r/2, title = "\n No pollinators", superunname = TRUE
             , line_col = "black")
mk_graph_3sp(alphamat = alpha_link, rs = r_link/2, title = "\n Link", superunname = TRUE
             , line_col = "DodgerBlue")
mk_graph_3sp(alphamat = alpha_no_link, rs = r_link/2, title = "\n No link", 
             superunname = TRUE, line_col = "darkorange")
#B) triangles
nopol <- print(simplex_plot(alpha, r, col = "black", line_col = "black"))
link <- print(simplex_plot(alpha_link, r_link, col = "DodgerBlue", line_col = "DodgerBlue"))
no_link <- print(simplex_plot(alpha_no_link, r_no_link, col = "darkorange", line_col = "darkorange"))
```

Next, we ask if the full community coexistence changes for the full community

```{r}
C <- diag(c(1,1,1,1,1,1), 6) # constraints

Omega_fit_no_link <- Omega_overlap(beta_no_link,C)*2
probs_no_link <- Probabilities_plant_pos(beta_no_link,6)

Omega_fit_link <- Omega_overlap(beta_link,C)*2
probs_link <- Probabilities_plant_pos(beta_link,6)

Omega_fit_analitical <- Omega_overlap(beta_no_link_analitical2,C)*2
probs_analitical <- Probabilities_plant_pos(beta_no_link_analitical2,6)

print(c(Omega_fit_no_link,probs_no_link))
print(c(Omega_fit_link,probs_link))
print(c(Omega_fit_analitical,probs_analitical))
#Those are probabilities of species coexsisting.
#Notice Omega's are very close... 

#plot
# barplot(as.matrix(rbind(probs_link, probs_analitical, probs_no_link)), beside = TRUE, 
#         las = 2, names.arg = c("Solanum", "Raphanus", "Vicea", "Osmia", 
#                                "Bombus", "Lucilia"), ylim = c(0,1), col = c("DodgerBlue",  
#                                                                             "orangered3",
#                                                                             "darkorange")
#         #legend.text = c("no link", "link", "anal. removed")
#         )

plot(x = 1,                 
     xlab = " ", ylab = "Probability",
     xlim = c(1, 12), ylim = c(0, 1),
     type = "n", las = 1)
segments(x0 = c(1,3,5,7,9,11), 
         y0 = probs_link, 
         x1 = c(2,4,6,8,10,12), 
         y1 = probs_analitical, 
         col = "orangered3")
segments(x0 = c(1,3,5,7,9,11), 
         y0 = probs_link, 
         x1 = c(2,4,6,8,10,12), 
         y1 = probs_no_link, 
         col = "darkorange")
points(x = c(1,3,5,7,9,11), y = probs_link, pch = 17, col = "DodgerBlue")
points(x = c(2,4,6,8,10,12), y = probs_analitical, pch = 15, col = "orangered3")
points(x = c(2,4,6,8,10,12), y = probs_no_link, pch = 19, col = "darkorange")

#in % #Figure S3.
change_analitical <- abs(100-(probs_analitical*100)/probs_link)
change_nolink <- abs(100-(probs_no_link*100)/probs_link)
plot(x = 1,                 
     xlab = " ", ylab = "absolute effect (%)",
     xlim = c(1, 12), ylim = c(0, 80),
     type = "n", las = 1)
segments(x0 = c(1,3,5,7,9,11), 
         y0 = change_analitical, 
         x1 = c(2,4,6,8,10,12), 
         y1 = change_nolink, 
         col = "grey")
points(x = c(1,3,5,7,9,11), y = change_analitical, pch = 15, col = "orangered3")
points(x = c(2,4,6,8,10,12), y = change_nolink, pch = 19, col = "darkorange")

segments(x0 = c(1,3,5,7,9,11), 
         y0 = c(-3,-3,-3,-3,-3,-3), 
         x1 = c(1,3,5,7,9,11), 
         y1 = change_analitical, 
         col = "orangered3")
segments(x0 = c(2,4,6,8,10,12), 
         y0 = c(-3,-3,-3,-3,-3,-3), 
         x1 = c(2,4,6,8,10,12), 
         y1 = change_nolink, 
         col = "darkorange")
wilcox.test(change_analitical, change_nolink, paired = TRUE, alternative = "two.sided")
mean(change_analitical)
mean(change_nolink)
mean(change_analitical)-mean(change_nolink) 
```

We add the null model results to see how random change affects directionality. With this, we plot Fig 3B:

```{r}
load("data/reorganization_random_nacho.RData")
dat <- dat2
head(dat)
print(original_diff <- probs_no_link - probs_link) 
print(mean_random_diff <- colMeans(dat) - probs_link) 

change_analitical <- (100-(probs_analitical*100)/probs_link)
change_nolink <- (100-(probs_no_link*100)/probs_link)
change_null <- (100-(colMeans(dat)*100)/probs_link)
plot(x = 1,                 
     xlab = " ", ylab = "effect (%)",
     xlim = c(1, 12), ylim = c(-100, 80),
     type = "n", las = 1)
segments(x0 = c(1,3,5,7,9,11), 
         y0 = change_analitical, 
         x1 = c(2,4,6,8,10,12), 
         y1 = change_nolink, 
         col = "grey")
segments(x0 = c(1,3,5,7,9,11), 
         y0 = change_analitical, 
         x1 = c(2,4,6,8,10,12), 
         y1 = change_null, 
         col = "grey")
points(x = c(1,3,5,7,9,11), y = change_analitical, pch = 15, col = "orangered3")
points(x = c(2,4,6,8,10,12), y = change_nolink, pch = 19, col = "darkorange")
points(x = c(2,4,6,8,10,12), y = change_null, pch = 1, col = "black")
points(x = c(2,4,6,8,10,12), y = change_null, pch = 16, col = "lightgrey")
# segments(x0 = c(1,3,5,7,9,11), 
#          y0 = c(0,0,0,0,0,0), 
#          x1 = c(1,3,5,7,9,11), 
#          y1 = change_analitical, 
#          col = "orangered3")
# segments(x0 = c(2,4,6,8,10,12), 
#          y0 = c(0,0,0,0,0,0), 
#          x1 = c(2,4,6,8,10,12), 
#          y1 = change_nolink, 
#          col = "darkorange")
# segments(x0 = c(2,4,6,8,10,12), 
#          y0 = c(0,0,0,0,0,0), 
#          x1 = c(2,4,6,8,10,12), 
#          y1 = change_null, 
#          col = "grey")
abline(h = 0, lty = 3)
```


To plot Figure S3:

```{r}
#order T           R             H          O            B            F
par(mfrow = c(2,3))
hist(dat[,1], las = 1, main = "Tomato", xlab = "simulated")
abline(v = probs_no_link[1], col = "grey", lwd = 3)
abline(v = probs_link[1], col = "DodgerBlue", lty = 2, lwd = 3)
#abline(v = probs_analitical[1], col = "green")
abline(v = mean(dat[,1]), col = "orangered3", lwd = 3)
hist(dat[,2], las = 1, main = "Radish", xlab = "simulated")
abline(v = probs_no_link[2], col = "grey", lwd = 3)
abline(v = probs_link[2], col = "DodgerBlue", lty = 2, lwd = 3)
#abline(v = probs_analitical[2], col = "green")
abline(v = mean(dat[,2]), col = "orangered3", lwd = 3)
hist(dat[,3], las = 1, main = "Field Bean", xlab = "simulated")
abline(v = probs_no_link[3], col = "grey", lwd = 3)
abline(v = probs_link[3], col = "DodgerBlue", lty = 2, lwd = 3)
#abline(v = probs_analitical[3], col = "green")
abline(v = mean(dat[,3]), col = "orangered3", lwd = 3)
hist(dat[,4], las = 1, main = "Mason bee", xlab = "simulated")
abline(v = probs_no_link[4], col = "grey", lwd = 3)
abline(v = probs_link[4], col = "DodgerBlue", lty = 2, lwd = 3)
#abline(v = probs_analitical[4], col = "green")
abline(v = mean(dat[,4]), col = "orangered3", lwd = 3)
hist(dat[,5], las = 1, main = "Bumblebee", xlab = "simulated")
abline(v = probs_no_link[5], col = "grey", lwd = 3)
abline(v = probs_link[5], col = "DodgerBlue", lty = 2, lwd = 3)
#abline(v = probs_analitical[5], col = "green")
abline(v = mean(dat[,5]), col = "orangered3", lwd = 3)
hist(dat[,6], las = 1, main = "Green bottle fly", xlab = "simulated")
abline(v = probs_no_link[6], col = "grey", lwd = 3)
abline(v = probs_link[6], col = "DodgerBlue", lty = 2, lwd = 3)
#abline(v = probs_analitical[6], col = "green")
abline(v = mean(dat[,6]), col = "orangered3", lwd = 3)
par(mfrow = c(1,1))

```

